---
title: "rcox-package example in 3D"
author: "Tuomas Rajala"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example 3D}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\newcommand{\x}{\mathbf{x}}


This doc gives a short example of how to simulate shot-noise and product-noise Cox patterns in 3D.

See the "Example_1" vignette for basics.


```{r, message=FALSE, warning=FALSE}
library(rcox)
```

# Simulation in 3D

Two examples:

```{r, fig.width=6}
set.seed(1) 
x1 <- matrix( runif( 10 * 3), ncol = 3 )
x2 <- cbind(0.5, 0.5, 0.5)
bb3 <- cbind(0:1, 0:1, 0:1) # bounding box for later
```

We set up the models like this:

```{r}
lam1 <- lambda(x1, kernel = "gauss", sigma = 0.05, alpha = 5, type = "sum")
lam2 <- lambda(x2, kernel = "step", sigma = 0.1, alpha = c(1, 100), type = "prod")
```

The documentation lists the options.
## Check the 3D field

One can evaluate the field at any point. Let's slice through the window with $z=0.5$ plane:

```{r}
# make a slice z=0.5
ng <- seq(0, 1, l = 50)
grid3 <- as.matrix( expand.grid( ng, ng , 0.5) )
field1 <- coxintensity2matrix(lam1, bbox = bb3, at = grid3)
field2 <- coxintensity2matrix(lam2, bbox = bb3, nx = 128, at = grid3)
```

```{r, fig.width=6}
par(mfrow=c(1,2))
plot3 <- function(f) image(matrix(f, ncol = length(ng)), x=ng, y=ng, asp=1) 
plot3(field1)
plot3(field2)
```

Red means low intensity (not preferred), white means high intensity (preferred). 

## Simulating

The simulation uses either a thinning approach (free point count) or Metropolis-Hastings (fixed point count). The calls:

```{r}
s1 <- rcox(lambda = lam1, bbox = bb3)
s2 <- rcox(lambda = lam2, bbox = bb3, n = 100)
# turn to ppp-objects
y1 <- cox2ppp(s1)
y2 <- cox2ppp(s2)
```

```{r, fig.width=6}
par(mfrow=c(1,2), mar=c(1,1,1,1))
plot(y1)
plot(y2)
```

There we have it.

*EOF*






